Do motif roles provide unique information about species risk of extinction?\\ \medskip Supplementary Information

### 2021-06-04 manuscript/supplement_motifs.tex
trying to sync git to overleaf

\section*{Table of contents}
    \begin{spacing}{1.0}

    \begin{enumerate}

        \item How do counts of motifs in a network profile vary with S and C? \\

            \begin{itemize}
            \item Same methods as main text except using counts instead of proportions to define motif profiles
            \end{itemize}

        \item Motif profiles differed slightly between network sets

           \begin{itemize}
                \item Comparing proportions of motifs in un-filtered, filtered, and acyclic networks.
                \item AFAIK not referred to in text
            \end{itemize}

         \item Persistence and count-based motif participation

           \begin{itemize}
                \item Should be as main text section 4, but counts
            \end{itemize}

        \item Motif-persistence relationships and network size
            \begin{itemize}
                \item Detailed results for variability of slopes for persistence $\approx$ motif participation in networks of different sizes.
                \item Main text refers to C, moved figure to Main while we check that we've got everything.
            \end{itemize}

        \item How does persistence vary with degree and trophic level?

            \begin{itemize}
                \item Heat map showing persistence vs. TL with BP=0
                \item Heat map showing persistence vs. TL with BP=1
            \end{itemize}

    \end{enumerate}
\end{spacing}
\clearpage

\section{How do counts of motifs in a network profile vary with S and C?}

    As expected, all four motifs had higher counts in larger and more highly connected networks.
    The precise effects of size, connectance, and their interaction differed between motifs  (Table~\ref{network_motif_lms}).
    This suggests that the meso-scale structure of a network varies with size and connectance, not just the total number of motifs.
    Indeed, the motif profile of a network is significantly related to its size, connectance, and their interaction ($F_{1,2996}$=2279, $p$\textless0.001; $F_{1,2996}$=3067, $p$\textless0.001; and $F_{1,2996}$=1187, $p$\textless0.001, respectively).

    \begin{table}[h!]
        \centering
        \caption{Coefficients for linear regressions relating the total count of all motifs or the count of each three-species motif in a network to network size, connectance, and their interaction. Coefficients in \textbf{bold} are significant (\textless0.05). Note that, while there are 13 unique three-species motifs, only four are possible in acyclic networks such as those considered here.}
       \label{network_motif_lms}
       \begin{tabular}{c|c c c c c}
        Motif & Intercept & Size & Connectance & Interaction \\
        \hline
        Total count & \textbf{-2.85$\times10^3$} & \textbf{46.6} & \textbf{-2.12$\times10^{5}$} & \textbf{4.72$\times10^{3}$} \\
        \hline
        Thee-species Chain & \textbf{-1.40$\times10^{3}$} & \textbf{26.1} & \textbf{-3.74$\times10^{4}$} & \textbf{8.58$\times10^{2}$} \\
        Apparent Competition & \textbf{-2.23$\times10^3$} & \textbf{42.0} & 
        \textbf{-7.26$\times10^4$} & \textbf{1.63$\times10^3$} \\
        Direct Competition & 26.6 & 0.777 & \textbf{-4.85$\times10^4$} & \textbf{9.89$\times10^2$} \\
        Omnivory & \textbf{7.57} & \textbf{-22.2} & \textbf{-5.37$\times10^4$} & \textbf{1.24$\times10^3$} \\
        \hline
        \end{tabular}
    \end{table}

    \begin{figure}[h!]
        \centering
        \includegraphics[width=.75\textwidth]{figures/motif_profile_dispersion.pdf}
        \caption{Dispersion of motif profiles within a network about the centroid for that combination of network size and connectance. Each circle represents a single network; circle colour indicates network size. Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
        \label{dispersion_rawmotifs}
    \end{figure}

    Motif profiles were similarly variable across networks with different sizes, but were not homogeneously variable across connectances or size-connectance combinations ($F_{5,2994}$=0.365, $p$=0.873; $F_{4,2995}$=4.49, $p$=0.001; and $F_{29,2970}$=27.1, $p$\textless0.001 , respectively).
    These differences in variability of motif profiles can cause false positives in the PERMANOVA.
    Variability in motif profiles was lowest among large networks with moderate connectance and was highest among the least-connected and most-connected networks we considered (Fig.~\ref{dispersion_rawmotifs}).

    \begin{figure}[h!]
        \centering
        \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
        \caption{The effect of number of times a species appears (x-axis) in any of the various motifs (columns) on persistence (y-axis). The effect of each motif participation is fitted by linear mixed-effect models, following Equation~\ref{absreq}, except that the disturbance levels are fitted separately. The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
        \label{fig:abs_lmer_all}
    \end{figure}

@@

\section{Motif profiles differed slightly in different network sets}

@@

\section{Persistence and count-based motif participation}

    \subsection*{Methods}

        We test whether participating in higher absolute numbers (i.e., counts) of a motif increases persistence using a set of models that parallels those used to test whether having a greater proportion of the role made up by a particular motif (see \emph{Main Text}).  
        Specifically, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
        In each case, we considered only non-basal species (i.e., those with at least one prey) as the persistence of basal species is determined only by their baseline probability.
        These models included the absolute number of times the focal species appeared in the focal motif, the increase in probability of extinction for basal species, and the interaction between the two as well as a random intercept for the interaction between species richness and connectance (equations.
        All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
        We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.

            The formula for these LMMs are:         
            \begin{equation}
                \Psi_{ik} \approx \chi_{i} + \pi_{disturbed_k} + \chi_{i}\pi_{disturbed_k} + S:C_{i} ,
                \label{absreq}
            \end{equation}

            \noindent where $\Psi_ik$ is the persistence of species $i$ during disturbance level $k$, $\chi_{i}$ is the number of times species $i$ appears in the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.

            % Additionally, we performed a similar linear regression of persistence against motif participation within each simulated network and for each level of disturbance to basal resources (XX regressions total).
            % These regressions were fit using the R~\citep{R} base function `lm'.
            % We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.

    \subsection*{Results}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
            \caption{The effect of number of times a species appear (x-axis) in any of the various motifs (columns) on persistence (y-axis). The effect of each motif participation is fitted by linear mixed-effect models, following \emph{Equation 2; Main Text}, except that the disturbance levels are fitted separately. The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
            \label{fig:abs_lmer_all}
        \end{figure}

>>>>

% \section*{Table of contents}
%     \begin{spacing}{1.0}

%     \begin{enumerate}

%         \item How do counts of motifs in a network profile vary with S and C? \\

%             \begin{itemize}
%             \item Same methods as main text except using counts instead of proportions to define motif profiles
%             \end{itemize}

%         \item Motif profiles differed slightly between network sets

%            \begin{itemize}
%                 \item Comparing proportions of motifs in un-filtered, filtered, and acyclic networks.
%                 \item AFAIK not referred to in text
%             \end{itemize}

%          \item Persistence and count-based motif participation

%            \begin{itemize}
%                 \item Should be as main text section 4, but counts
%             \end{itemize}

%         \item Motif-persistence relationships and network size
%             \begin{itemize}
%                 \item Detailed results for variability of slopes for persistence $\approx$ motif participation in networks of different sizes.
%                 \item Main text refers to C, moved figure to Main while we check that we've got everything.
%             \end{itemize}

%         \item How does persistence vary with degree and trophic level?

%             \begin{itemize}
%                 \item Heat map showing persistence vs. TL with BP=0
%                 \item Heat map showing persistence vs. TL with BP=1
%             \end{itemize}

%     \end{enumerate}
% \end{spacing}
% \clearpage

\section{SA: Persistence and global network properties}

    Connectance and species richness had smaller effects on consumer persistence than disturbance (Table~\ref{tab:per_vs_SC}, Fig.~\ref{fig:heatmap_CS}).

    \begin{figure}[h!]
      \includegraphics[width=0.95\textwidth]{figures/heatmap_CS_BPcompare.eps}
       \caption{Heat maps of average consumer persistence in networks with varying connectance (y-axis) and network size (x-axis) at two levels of disturbance to basal resources. Persistence decreases with darker color in the heat map. When basal species are not disturbed, all species experience a baseline extinction risk of $\pi = 0.1$. When basal species are highly disturbed, their baseline extinction risk rises to $\pi = 0.5$ while the baseline risk of consumers is not changed.}
       \label{fig:heatmap_CS}
    \end{figure}

    \begin{table}[hb!]
        \caption{Persistence of consumers within a network depends on network size (S) and connectance (C), the level of disturbance to basal resources (D), and interactions between them. Here we show the effect sizes ($\beta$) and $p$-values for the regression described above. All predictors were centered and scaled before fitting the model. The level of disturbance corresponds to basal resources having a probability of extinction ranging between $0.1 - 0.5$ (see \emph{Methods}).
        }
        \label{tab:per_vs_SC}
        \centering
        \begin{tabular}{c|c c |}
            Predictor & $\beta$ & $p$-value \\
            \hline
            Intercept & 0.572 & \textless0.001 \\
            Size & -5.15$\times10^{-4}$  & \textless0.001 \\
            Connectance & -0.0129 & \textless0.001 \\
            Disturbance & -0.139 & \textless0.001 \\
            S:C & -5.24$\times10^{-4}$ & \textless0.001 \\
            S:D & -3.48$\times10^{-3}$ & \textless0.001 \\
            C:D & -7.00$\times10^{-5}$ & 0.095 \\
            S:C:D: & -2.69$\times10^{-4}$ & \textless0.001 \\
        \end{tabular}
    \end{table}

@@

\section{SB: Motif profiles and global properties}

    Normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).These differences in variability can cause false positives in the PERMANOVA test. Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).

   \begin{figure}[h!]
       \centering
       \includegraphics[width=.75\textwidth]{figures/proportion_profile_dispersion.eps}
       \caption{Dispersion of normalized motif profiles (i.e., proportions of each motif) within a network about the centroid for that combination of network size and connectance. Each circle represents a single network; circle colour indicates network size. Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
       \label{dispersion_normmotifs}
    \end{figure}

    The frequencies of the four possible motifs were related to global network properties.

    \begin{table}[hb!]
        \centering
        \caption{Coefficients for linear regressions relating the proportion of each motif in the total motif profile of a network to network size, connectance, and their interaction. Coefficients in \textbf{bold} are significant (\textless0.05).}
       \label{network_prop_lms}
       \begin{tabular}{c|c c c c c}
            Motif & Intercept & Size & Connectance & Interaction \\
            \hline
            Three-species Chain & \textbf{0.262} & \textbf{1.78$\times10^{-4}$} & \textbf{-0.157} & \textbf{-3.27$\times10^{-3}$} \\
            Apparent Competition & \textbf{0.541} & \textbf{-5.06$\times10^{-4}$} & \textbf{-1.00} & \textbf{2.59$\times10^{-3}$} \\
            Direct Competition & \textbf{0.193} & -3.56$\times10^{-5}$ & \textbf{-0.556} & \textbf{5.92$\times10^{-3}$} \\
            Omnivory & 4.82$\times10^{-3}$ & \textbf{3.64$\times10^{-4}$} & \textbf{1.71} & \textbf{-5.24$\times10^{-3}$}\\
            \hline
            \end{tabular}
    \end{table}        
\clearpage

\section{SC: Persistence and motif profiles}

    Persistence varied with the frequencies of different motifs in a network profile.

    \begin{table}[h!]
        \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction. Predictors were centered and scaled before model fitting. All effects were significant ($p$\textless0.001). We also provide the $R^2$. }
        \label{motif_profile_tab}
        \centering
        \begin{tabular}{c|c c c c c | c}
            Motif & Intercept & Proportion & Disturbance & Interaction & $R^2$\\
            \hline
            Omnivory & 0.572 & -0.0133 & -0.139 & -4.89$\times10^{-4}$ & 0.908 \\
            Apparent competition & 0.572 & 0.0105 & -0.139 & 4.78$\times10^{-4}$ & 0.905 \\
            Direct competition & 0.572 & 2.62$\times10^{-3}$ & -0.139 & 3.84$\times10^{-4}$ & 0.901 \\
            Three-species chain & 0.572 & 6.12$\times10^{-3}$ & -0.139 & -1.90$\times10^{-4}$ & 0.901 \\ 
        \end{tabular}
    \end{table}
\clearpage

\section{SD: Persistence and motif participation}

    Persistence varied with motif participation.

    \begin{table}[h!]
        \centering
        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}). Predictors were centered and scaled before model fitting. We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
        \label{tab:proportion}                \footnotesize
        \begin{tabular}{c|c c | c c | c c | c c | c c |}
        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
        \hline
        3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
        App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
        Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
        Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
        \end{tabular}
    \end{table}        
\clearpage

\section{SE: Motif-persistence relationships \& network size}

    \begin{figure}[h!]
         \centering
         % \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
        \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. The slopes are derived from linear mixed-effect models, with a random intercept for the interaction between species richness and network size. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence. The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend). Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right). Rows show various levels of network size. The vertical line indicate zero on the x-axis.}
        \label{fig:density_prop_S}
    \end{figure}

    For the apparent and direct competition motifs, the LMMs include a positive interaction between disturbance and the proportion of the motif in a species' role (Fig. 3 and Table 1; \emph{Main Text}).
    This means that having more of the role made up of these two motifs is increasingly beneficial at higher levels of disturbance, a trend which is also evident from the increasing fraction of positive slopes with increasing disturbance based on LRs of each network at each level of disturbance. % (Figure~\ref{fig:density_prop_S}).
    As with the trends for omnivory and three-species chains, these increases are strongest in large networks. % (bottom row, Figure~\ref{fig:density_prop_S}).
    % For direct competition, the fraction of slopes are similar across disturbance levels. Although this interaction is not positive when accounting for network size, the fraction of positive slopes are throughout high.
    The increase in the proportion of positive slopes was much stronger for apparent competition than direct competition, consistent with the smaller (but still significant) interaction term for between direct competition in the LMMs.
    A high proportion of direct competition in a species' role tended to be associated with greater probability of persistence across levels of disturbance and network size.
    % In conclusion, species richness does not considerably alter the trends visible in the LMMs. High species richness more clearly display an interaction between motif proportion, persistences and disturbance levels, while this effect is smaller with low species richness. 
\clearpage

\section{SF: Persistence, degree, and trophic level}

    Persistence varies with in-degree, trophic level, and their interaction.

    \begin{table}[h!]
        \caption{Coeffcients ($\beta$) for regressions of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction. Predictors were centered and scaled before model fitting. All $p$-values were \textless0.001. The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.}
        \label{tab:per_vs_TLdeg}
        \centering
        \begin{tabular}{l|c  c  c |}
            Simple role measure & Role & Disturbance & Interaction \\
            \hline
            STL & -0.0218 & -0.139 & 0.00508 \\
            in-degree & -0.00130 & -0.139 & -0.00908 \\
        \end{tabular}
    \end{table}

    \begin{figure}[h!]
     \centering
     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
     \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$. The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis). Persistence decreases with darker color in the heat map.}
     \label{fig:heatmap_stl_BP0}
    \end{figure}

    \begin{figure}[h!]
     \centering
     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
     \caption{Primary producers experience a disturbance reducing their probability to go extinct to $\pi_{disturbed} = 0.5$ All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$. The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis). Persistence decreases with darker color in the heat map.}
     \label{fig:heatmap_stl_BP1}
    \end{figure}
\clearpage

\section{SG: Motif participation and network properties}

    Motif participation varies with global-scale network properies.

    \begin{table}[hb!]
        \centering
        \caption{Motif participation varies with size and connectance. Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors. 
        We also show the effects ($\beta$) of network size, connectance, and their interaction in linear models predicting the proportion of each motif in a participation vector.
        All effects were significant ($p$\textless0.001).
        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
        \label{tab:partic_vs_SC}   
        \footnotesize
        \begin{tabular}{c|c c c | c c c}
            Motif & Min & Max & Mean (SD) & Size & Connectance & Interaction \\
            \hline
            3-sp Chain & 0 & 1 & 0.249 (0.127) & 2.22$\times10^{-4}$ & -0.131 & -3.20$\times10^{-3}$ \\
            App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.06$\times10^{-4}$ & -1.00 & 2.59$\times10^{-3}$ \\
            Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -3.56$\times10^{-5}$ & -0.556 & 5.92$\times10^{-3}$ \\
            Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.64$\times10^{-4}$ & 1.71 & -5.24$\times10^{-3}$ \\   
            \hline
            \end{tabular}
            \end{table}

    \begin{figure}[h!]
        \centering
        \includegraphics[width=.95\textwidth]{figures/roles_vs_SC.eps}

        \caption{The proportion of each motif in a species' participation vector varied with network size, connectance, and their interaction. These relationships were steepest for apparent competition and omnivory.}
        \label{fig:roles_vs_SC}
    \end{figure}
\clearpage

\section*{SH: Motif roles and network properties}

    \subsection*{Degree and trophic level}

        \begin{table}[h!]
            \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against in-degree, STL, and their interaction.}
            \label{DT_tab}
            \footnotesize
            \begin{tabular}{c c | c c | cc | cc |}
            Motif & Position & \multicolumn{2}{c}{Degree} & \multicolumn{2}{c}{STL} & \multicolumn{2}{c}{Interaction} \\
            & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
            \hline
            Chain & bottom & -1.47 $\times10^{-5}$ & 0.769 & -0.0218 & \textless0.001 & -0.00160 & \textless0.001 \\
            Chain & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & 0.275 \\
            Chain & top & -0.0172 & \textless0.001 & 0.0259 & \textless0.001 & 0.00919 & \textless0.001 \\
            \hline
            Omnivory & bottom & 6.14$\times10^{-4}$ & \textless0.001 & -0.00270 & \textless0.001 & -0.00112 & \textless0.001 \\
            Omnivory & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & \textless0.001 \\
            Omnivory & top & 0.00413 & \textless0.001 & 0.00381 & \textless0.001 & 9.96$\times10^{-5}$ & \textless0.001 \\
            \hline
            App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
            App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
            \hline
            Dir. comp. & bottom & 8.12$\times10^{-4}$ & \textless0.001 & -5.57$\times10^{-3}$ & \textless0.001 & -1.37$\times10^{-3}$ & \textless0.001 \\
            Dir. comp. & top & -0.00405 & \textless0.001 & -0.0435 & \textless0.001 & 0.00260 & \textless0.001 \\
            \hline
            \end{tabular}
            \end{table}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=.75\textwidth]{figures/positions_byTL_Omnivory.eps}
            \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
            \label{omnivory_DT}
        \end{figure}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=.75\textwidth]{figures/positions_byTL_Apparentcompetition.eps}
            \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
            \label{appcomp_DT}
        \end{figure}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=.75\textwidth]{figures/positions_byTL_Directcompetition.eps}
            \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
            \label{dircomp_DT}
        \end{figure}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=.75\textwidth]{figures/positions_byTL_Three-specieschain.eps}
            \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
            \label{appcomp_DT}
        \end{figure}

    \subsection*{Network size and connectance}

        \begin{table}[h!]
            \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against network size, connectance, and their interaction.}
            \label{DT_tab}
            \footnotesize
            \begin{tabular}{c c | c c | cc | cc |}
            Motif & Position & \multicolumn{2}{c}{Size} & \multicolumn{2}{c}{Connectance} & \multicolumn{2}{c}{Interaction} \\
            & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
            \hline
            Chain & bottom & -4.63$\times10^{-5}$ & \textless0.001 & -0.108 & \textless0.001 & -1.94$\times10^{-4}$ & 0.035 \\
            Chain & middle & 1.68$\times10^{-4}$ & \textless0.001 & -0.0530 & \textless0.001 & -0.00212 & \textless0.001 \\
            Chain & top & 1.00$\times10^{-4}$ & \textless0.001 & 0.0304 & 0.015 & -8.79$\times10^{-4}$ & \textless0.001 \\
            \hline
            Omnivory & bottom & 2.00$\times10^{-4}$ & \textless0.001 & 0.575 & \textless0.001 & -0.00240 & \textless0.001 \\
            Omnivory & middle & 1.32$\times10^{-4}$ & \textless0.001 & 0.551 & \textless0.001 & -0.00215 & \textless0.001 \\
            Omnivory & top & 6.42$\times10^{-5}$ & \textless0.001 & 0.448 & \textless0.001 & -0.00115 & \textless0.001 \\
            \hline
            App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
            App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
            \hline
            Dir. comp. & bottom & 7.94$\times10^{-5}$ & \textless0.001 & -0.0663 & \textless0.001 & 0.00101 & \textless0.001 \\
            Dir. comp. & top & -2.45$\times10^{-4}$ & \textless0.001 & -0.602 & \textless0.001 & 0.00610 & \textless0.001 \\
            \hline
            \end{tabular}
            \end{table}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=.75\textwidth]{figures/positions_bySC_Omnivory.eps}
            \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
            \label{omnivory_DT}
        \end{figure}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=.75\textwidth]{figures/positions_bySC_Apparentcompetition.eps}
            \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
            \label{appcomp_DT}
        \end{figure}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=.75\textwidth]{figures/positions_bySC_Directcompetition.eps}
            \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
            \label{dircomp_DT}
        \end{figure}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=.75\textwidth]{figures/positions_bySC_Three-specieschain.eps}
            \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
            \label{appcomp_DT}
        \end{figure}

% \section{How do counts of motifs in a network profile vary with S and C?}

%     As expected, all four motifs had higher counts in larger and more highly connected networks.
%     The precise effects of size, connectance, and their interaction differed between motifs  (Table~\ref{network_motif_lms}).
%     This suggests that the meso-scale structure of a network varies with size and connectance, not just the total number of motifs.
%     Indeed, the motif profile of a network is significantly related to its size, connectance, and their interaction ($F_{1,2996}$=2279, $p$\textless0.001; $F_{1,2996}$=3067, $p$\textless0.001; and $F_{1,2996}$=1187, $p$\textless0.001, respectively).

%     \begin{table}[h!]
%         \centering
%         \caption{Coefficients for linear regressions relating the total count of all motifs or the count of each three-species motif in a network to network size, connectance, and their interaction. Coefficients in \textbf{bold} are significant (\textless0.05). Note that, while there are 13 unique three-species motifs, only four are possible in acyclic networks such as those considered here.}
%        \label{network_motif_lms}
%        \begin{tabular}{c|c c c c c}
%         Motif & Intercept & Size & Connectance & Interaction \\
%         \hline
%         Total count & \textbf{-2.85$\times10^3$} & \textbf{46.6} & \textbf{-2.12$\times10^{5}$} & \textbf{4.72$\times10^{3}$} \\
%         \hline
%         Thee-species Chain & \textbf{-1.40$\times10^{3}$} & \textbf{26.1} & \textbf{-3.74$\times10^{4}$} & \textbf{8.58$\times10^{2}$} \\
%         Apparent Competition & \textbf{-2.23$\times10^3$} & \textbf{42.0} & 
%         \textbf{-7.26$\times10^4$} & \textbf{1.63$\times10^3$} \\
%         Direct Competition & 26.6 & 0.777 & \textbf{-4.85$\times10^4$} & \textbf{9.89$\times10^2$} \\
%         Omnivory & \textbf{7.57} & \textbf{-22.2} & \textbf{-5.37$\times10^4$} & \textbf{1.24$\times10^3$} \\
%         \hline
%         \end{tabular}
%     \end{table}

%     \begin{figure}[h!]
%         \centering
%         \includegraphics[width=.75\textwidth]{figures/motif_profile_dispersion.pdf}
%         \caption{Dispersion of motif profiles within a network about the centroid for that combination of network size and connectance. Each circle represents a single network; circle colour indicates network size. Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
%         \label{dispersion_rawmotifs}
%     \end{figure}

%     Motif profiles were similarly variable across networks with different sizes, but were not homogeneously variable across connectances or size-connectance combinations ($F_{5,2994}$=0.365, $p$=0.873; $F_{4,2995}$=4.49, $p$=0.001; and $F_{29,2970}$=27.1, $p$\textless0.001 , respectively).
%     These differences in variability of motif profiles can cause false positives in the PERMANOVA.
%     Variability in motif profiles was lowest among large networks with moderate connectance and was highest among the least-connected and most-connected networks we considered (Fig.~\ref{dispersion_rawmotifs}).

%     \begin{figure}[h!]
%         \centering
%         \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
%         \caption{The effect of number of times a species appears (x-axis) in any of the various motifs (columns) on persistence (y-axis). The effect of each motif participation is fitted by linear mixed-effect models, following Equation~\ref{absreq}, except that the disturbance levels are fitted separately. The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
%         \label{fig:abs_lmer_all}
%     \end{figure}

\section{Motif profiles in different network sets}

@@

    \begin{table}[h!]
        \centering
        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}). Predictors were centered and scaled before model fitting. We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
        \label{tab:proportion}                \footnotesize
        \begin{tabular}{c|c c | c c | c c | c c | c c |}
        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
        \hline
        3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
        App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
        Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
        Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
        \end{tabular}
    \end{table}        
\clearpage

% \section{Persistence and count-based motif participation}

%     \subsection*{Methods}

%         We test whether participating in higher absolute numbers (i.e., counts) of a motif increases persistence using a set of models that parallels those used to test whether having a greater proportion of the role made up by a particular motif (see \emph{Main Text}).  
%         Specifically, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
%         In each case, we considered only non-basal species (i.e., those with at least one prey) as the persistence of basal species is determined only by their baseline probability.
%         These models included the absolute number of times the focal species appeared in the focal motif, the increase in probability of extinction for basal species, and the interaction between the two as well as a random intercept for the interaction between species richness and connectance (equations.
%         All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
%         We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.

%             The formula for these LMMs are:         
%             \begin{equation}
%                 \Psi_{ik} \approx \chi_{i} + \pi_{disturbed_k} + \chi_{i}\pi_{disturbed_k} + S:C_{i} ,
%                 \label{absreq}
%             \end{equation}

%             \noindent where $\Psi_ik$ is the persistence of species $i$ during disturbance level $k$, $\chi_{i}$ is the number of times species $i$ appears in the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
